---
title: "Electric power sector transition scenarios"
author: '@ideea'
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Configuration
```{r, eval = T}
nreg <- 7 # number of regions
offshore <- FALSE # TRUE/FALSE
islands <- FALSE
# time resolution
base_year <- 2020
ndays_sample <- 12 # 1 day per month, 12 months
# ndays_sample <- 36 # 3 days per month, 12 months
# ndays_sample <- 60 # 5 days per month, 12 months
# ndays_sample <- 84 # 7 days per month, 12 months
# ndays_sample <- 365 # full year

```

```{r, eval=TRUE, include=FALSE, message=FALSE, warning=FALSE}
solver_gams <- list(
  cmdline = "/Applications/GAMS25.1/sysdir/gams energyRt.gms", # MacOS/Linux
  # cmdline = "C:/GAMS/win64/32.2/gams.exe energyRt.gms", # Windows
  lang = "GAMS",
  solver = "cplex"
  # solver = "cbc"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(IDEEA)
library(sf)
library(glue)
# remotes::install_github("energyRt/timeslices")

regN <- glue("reg", nreg)
regN_off <- paste0(regN, "_off")

```

## Regions

```{r regions}
gis_sf <- get_ideea_map(nreg, offshore = T, reg_off = T)
plot(gis_sf["reg_off"], key.width = lcm(4))

# aggregating data (!!!write a function)
agg36toN <- get_ideea_map(46, offshore = T, rename = F, islands = T) %>% 
  st_drop_geometry() %>%
  select(reg36, all_of(regN), offshore) %>% unique() %>%
  mutate(reg36_off = if_else(offshore, paste0(reg36, "_off"), reg36))

agg36toN[[regN_off]] <- if_else(agg36toN$offshore, 
                                paste0(agg36toN[[regN]], "_off"), 
                                agg36toN[[regN]])

```

## Time resolution  

```{r}
# full 24x365 sets
tsl_levels <- ideea_modules$time_scales$tsl_levels
tsl_sets <- ideea_modules$time_scales$tsl_sets

# subsets
if (ndays_sample == 365) {
  # full set (8760h)
  YDAY_SUBSET <- tsl_levels$YDAY
  SLICE_SUBSET <- tsl_sets$slice
} else if (ndays_sample == 12) {
  # 12 representative days, mid-months
  YDAY_SUBSET <- ideea_modules$time_scales$yday_1day_x_12mon
  # time-slices levels
  tsl_levels$YDAY <- YDAY_SUBSET
  # time-slices sets, full table
  tsl_sets <- tsl_sets %>% filter(YDAY %in% YDAY_SUBSET)
  SLICE_SUBSET <- tsl_sets$slice
} else if (ndays_sample == 36) {
  # 12 representative days, mid-months
  YDAY_SUBSET <- ideea_modules$time_scales$yday_3day_x_12mon
  # time-slices levels
  tsl_levels$YDAY <- YDAY_SUBSET
  # time-slices sets, full table
  tsl_sets <- tsl_sets %>% filter(YDAY %in% YDAY_SUBSET)
  SLICE_SUBSET <- tsl_sets$slice
} else if (ndays_sample == 60) {
  # 12 representative days, mid-months
  YDAY_SUBSET <- ideea_modules$time_scales$yday_5day_x_12mon
  # time-slices levels
  tsl_levels$YDAY <- YDAY_SUBSET
  # time-slices sets, full table
  tsl_sets <- tsl_sets %>% filter(YDAY %in% YDAY_SUBSET)
  SLICE_SUBSET <- tsl_sets$slice
} else if (ndays_sample == 84) {
  # 12 representative days, mid-months
  YDAY_SUBSET <- ideea_modules$time_scales$yday_7day_x_12mon
  # time-slices levels
  tsl_levels$YDAY <- YDAY_SUBSET
  # time-slices sets, full table
  tsl_sets <- tsl_sets %>% filter(YDAY %in% YDAY_SUBSET)
  SLICE_SUBSET <- tsl_sets$slice
} else {
  stop("Unsupported value in 'ndays_sample': ", ndays_sample)
}
# 
SUBSET_HOURS <- length(SLICE_SUBSET) # total hours in the subset
FRACT_YEAR <- SUBSET_HOURS / 8760    # fraction in a year
print("Subannual time resolution:")
print(paste("   Total number of days a year:", length(YDAY_SUBSET)))
print(paste("   Total number of hours per year:", SUBSET_HOURS))

```


## Commodities
```{r}
# energy
ELC <- newCommodity(
  name = "ELC", 
  description = "Electricity",
  unit = "GWh", 
  slice = "HOUR"
)

# # using saved data from `energy` module
# COA <- ideea_modules$energy$COA
# GAS <- ideea_modules$energy$GAS
# BIO <- ideea_modules$energy$BIO

# emissions
CO2 <- newCommodity(
  name = "CO2", 
  description = "Carbon emissions",
  unit = "kt",
  slice = "ANNUAL"
)

NOX <- newCommodity(
  name = "NOX", 
  description = "Nitrogen oxide emissions NOx",
  unit = "kt",
  slice = "ANNUAL"
)

SOX <- newCommodity(
  name = "SOX", 
  description = "Sulfur oxide emissions SOx",
  unit = "kt",
  slice = "ANNUAL"
)

PM <- newCommodity(
  name = "PM", 
  description = "Particulate matter (particle pollution)",
  unit = "kt",
  slice = "ANNUAL"
)

REN <- newCommodity(
  name = "REN",
  description = "Generic renewable energy",
  unit = "GWh",
  slice = "ANNUAL"
)

repo_comm <- newRepository(name = "repo_comm", 
                           description = "Electricity & emissions commodity") %>%
  add(ELC, CO2, NOX, SOX, PM)

```


## Demand options

### Final/exogenous demand
```{r}
load_BY <- ideea_data$load_2019_MWh %>%
  # filter(slice %in% SLICE_SUBSET) %>%
  select(all_of(c("reg36", "slice", "MWh"))) %>%
  left_join(filter(agg36toN, !offshore), by = c(reg36 = "reg36")) %>%
  rename(reg_off = ends_with(regN_off), region = ends_with(regN)) %>%
  mutate(year = base_year) %>%
  group_by(region, reg_off, year, slice) %>%
  summarise(
    GWh = sum(MWh, na.rm = T)/1e3, # -> GWh
    .groups = "drop"
  ) 

DEMELC_BY <- newDemand(
  name = "DEMELC_BY",
  description = "houlry electric demand by region",
  commodity = "ELC",
  unit = ELC@unit,
  dem = data.frame(
    region = load_BY$region,
    # year = load_BY$year,
    slice = load_BY$slice, # comment to assign to all hours
    dem = load_BY$GWh # * dem_adj
  )
)

# load_0 <- load_BY %>% 
load_2x_2050 <-  
  rbind(
    mutate(load_BY, year = 2020, GWh = 0.0 * GWh),
    mutate(load_BY, year = 2050, GWh = 2 * GWh)
  )

load_BY %>% as.data.table()

DEMELC_2X <- newDemand(
  name = "DEMELC_2X",
  description = "Additional demand growth, proportional current load",
  commodity = "ELC",
  unit = ELC@unit,
  dem = data.frame(
    region = load_2x_2050$region,
    year = load_2x_2050$year,
    slice = load_2x_2050$slice, # comment to assign to all hours
    dem = load_2x_2050$GWh # * dem_adj
  )
)

```


### Demand-side technologies (groups) 
tbc
```{r}


```

## Supply & resources

### Energy
domestic reserves, mining, costs, import/export
```{r}
repo_supply <- newRepository(
  name = "repo_supply", 
  description = "Energy supply (markets)") %>%
  add(# from energy module
    # coal
    ideea_modules$energy$COA, # commodity
    ideea_modules$energy$SUP_COA, # supply
    ideea_modules$energy$IMP_COA,  # import 
    # natural gas
    ideea_modules$energy$GAS,
    ideea_modules$energy$SUP_GAS,
    ideea_modules$energy$IMP_GAS,
    # oil
    ideea_modules$energy$OIL,
    ideea_modules$energy$SUP_OIL,
    ideea_modules$energy$IMP_OIL,
    # biomass
    ideea_modules$energy$BIO,
    ideea_modules$energy$RES_BIO, # domestic resource of biomass
    # nuclear
    ideea_modules$energy$NUC,
    ideea_modules$energy$IMP_NUC, # Nuclear fuel international market
    # 
    REN, # all type of non-fuel (renewable) energy
    newSupply("RES_REN", commodity = "REN"),
    overwrite = F
  ) 

```

### Carbon sink / aquifers
potential CO2 storage, volume, locations

## Capacity factors

### Solar
NREL, MERRA2, (ERA5), potential location
```{r}

merra_solar <- ideea_data$merra_solar_cf_reg7 %>%
  # filter(slice %in% SLICE_SUBSET) %>%
  rename(region = reg7_off) %>%
  filter(region %in% gis_sf$region)

WSOL <- newWeather(
  name = "WSOL",
  description = "Solar capacity factors, tilted tracking system (tl)",
  region = unique(merra_solar$region),
  slice = "HOUR",
  weather = data.frame(
    region = merra_solar$region,
    slice = merra_solar$slice,
    # year = NA # all years
    wval = merra_solar$scf_tl
  )
)

```

### Wind
NREL, MERRA2, (ERA5), potential location

```{r}

merra_wind <- ideea_data$merra_wind_cf_reg7 %>%
  # filter(slice %in% SLICE_SUBSET) %>%
  rename(region = reg7_off) %>%
  filter(region %in% gis_sf$region)

WWIN <- newWeather(
  name = "WWIN",
  description = "Onshore wind 100m height",
  region = unique(merra_wind$region),
  slice = "HOUR",
  weather = data.frame(
    region = merra_wind$region,
    slice = merra_wind$slice,
    # year = NA # same for all years
    wval = merra_wind$wcf100m
  )
)

```

### Hydro
current capacity, seasonality, day/night 
```{r}

```


## Generating technologies
description of total capacity, by region, map

```{r}
cap <- ideea_data$generators %>%
  # left_join(ideea_data$reg_tbl) %>%
  group_by(across(all_of(regN)), offshore, mainland, primary_fuel) %>%
  summarize(
    capacity_mw = sum(capacity_mw, na.rm = T),
    .groups = "drop"
  ) %>%
  mutate(region = .data[[regN]], .before = 1) %>%
  select(-all_of(regN))
# ideea_r36_sf <- get_ideea_map(36, T, T)
# gis_sf

cap_sf <- gis_sf %>%
  right_join(cap) %>%
  filter(!is.na(primary_fuel))
  
ggplot() +
  geom_sf(data = gis_sf, fill = "grey") +
  geom_sf(aes(fill = capacity_mw/1e3), data = cap_sf) +
  scale_fill_viridis_c(option = "H", name = "GW", trans = "log10") +
  facet_wrap(~primary_fuel) +
  theme_bw() +
  theme(
    #panel.background = element_rect(fill = "aliceblue"),
    # panel.grid = element_line(color = "white", size = 0.8),
    axis.ticks = element_blank(),
    axis.text = element_blank())
ggsave("tmp/wri_installed_capacity.png", width = 6, height = 7, 
       scale = 1.25)

cap$primary_fuel %>% unique()

```


### Coal-fired
WRI dataset, new - the same and BAT

```{r}
cap_coa <- cap %>%
  filter(grepl("Coal", primary_fuel)) %>%
  filter(!offshore) %>%
  mutate(year = 2020, .before = 1)

# assume retirement of 2020 capacity by 2030
cap_coa_2030 <- cap_coa %>%
  mutate(year = 2030, capacity_mw = 0.8 * capacity_mw)

# assume retirement of 2020 capacity by 2030
cap_coa_2040 <- cap_coa %>%
  mutate(year = 2040, capacity_mw = 0.1 * capacity_mw)

# combine
cap_coa <- cap_coa %>% 
  bind_rows(cap_coa_2030) %>%
  bind_rows(cap_coa_2040)
# Note: the existing capacity will be linearly interpolated from 2020 to 2040
# cap = 0 after 2040

# load coal technology (assume Super-critical for all existing capacity)
ECOASUP_2020 <- ideea_modules$techs$ECOASUP@data$ECOASUP_2020
class(ECOASUP_2020) 

# update base-year technology with preexisting capacity
ECOASUP_2020 <- ECOASUP_2020 %>%
  update(stock = data.frame(
    region = cap_coa$region,
    year = cap_coa$year,
    stock = cap_coa$capacity_mw / 1e3 # in GW
  ))

# load most advanced coal techs for new investment
ECOA <- ideea_modules$techs$ECOAULT %>% # ultra-super-critical 
  add(ECOASUP_2020) # add tech with existing capacity
names(ECOA@data)

draw(ECOA@data$ECOASUP_2020)
draw(ECOA@data$ECOAULT_2040)

```

### Natural gas
```{r cap_gas}
cap_gas <- cap %>%
  filter(grepl("Gas", primary_fuel)) %>%
  filter(!offshore) %>%
  mutate(year = 2020, .before = 1)

# assume retirement of 2020 capacity by 2030
cap_gas_2030 <- cap_gas %>%
  mutate(year = 2030, capacity_mw = 0.8 * capacity_mw)

# assume retirement of 2020 capacity by 2030
cap_gas_2040 <- cap_gas %>%
  mutate(year = 2040, capacity_mw = 0.1 * capacity_mw)

# combine
cap_gas <- cap_gas %>% 
  bind_rows(cap_gas_2030) %>%
  bind_rows(cap_gas_2040)
# Note: the existing capacity will be linearly interpolated from 2020 to 2040
# cap = 0 after 2040

# load coal technology (assume Super-critical for all existing capacity)
ENGCC_2020 <- ideea_modules$techs$ENGCC@data$ENGCC_2020
class(ENGCC_2020) 

# update base-year technology with preexisting capacity
ENGCC_2020 <- ENGCC_2020 %>%
  update(stock = data.frame(
    region = cap_gas$region,
    year = cap_gas$year,
    stock = cap_gas$capacity_mw / 1e3 # in GW
  ))

# load most advanced coal techs for new investment
EGAS <- ideea_modules$techs$ENGCC %>% # ultra-super-critical 
  add(ENGCC_2020, overwrite = T) # add tech with existing capacity
names(EGAS@data)

draw(EGAS@data$ENGCC_2020)
# draw(ECOA@data$ENGCC_2040)

```

### Solar
existing - from WRI dataset
```{r}
# solar capacity in ~2020
cap_sol <- cap %>%
  filter(grepl("Solar", primary_fuel)) %>%
  mutate(year = 2020, .before = 1)

# assume retirement of 2020 capacity by 2030
cap_sol_2030 <- cap_sol %>%
  mutate(year = 2030, capacity_mw = 0.8 * capacity_mw)

# assume further retirement by 2040
cap_sol_2040 <- cap_sol %>%
  mutate(year = 2040, capacity_mw = 0.1 * capacity_mw)

# combine
cap_sol <- cap_sol %>% 
  bind_rows(cap_sol_2030) %>%
  bind_rows(cap_sol_2040)
# Note: the existing capacity will be linearly interpolated from 2020 to 2040
# cap = 0 after 2040

# load repository with solar technologies
ESOL <- ideea_modules$techs$ESOL
class(ESOL) # repository with solar techs
names(ESOL)

# update base-year technology with preexisting capacity
ESOL@data$ESOL_2020 <- ESOL@data$ESOL_2020 %>%
  update(stock = data.frame(
    region = cap_sol$region,
    year = cap_sol$year,
    stock = cap_sol$capacity_mw / 1e3 # in GW
  ))

# add weather factor name and parameter for each technology
ESOL@data <- lapply(ESOL@data, function(tech) update(
  tech,
  weather = data.frame(weather = "WSOL", waf.fx = 1),
  input = list(comm = "REN", combustion = 0)
))
names(ESOL@data)
ESOL@data$ESOL_2020@weather

ESOL@data$ESOL_2020 %>% draw()
```

### Wind
existing - from WRI dataset
```{r}
cap_win <- cap %>%
  filter(grepl("Wind", primary_fuel)) %>%
  mutate(year = 2020, .before = 1)

# assume retirement of 2020 capacity by 2030
cap_win_2030 <- cap_win %>%
  mutate(year = 2030, capacity_mw = 0.8 * capacity_mw)

# assume further retirement by 2040
cap_win_2040 <- cap_win %>%
  mutate(year = 2040, capacity_mw = 0.1 * capacity_mw)

# combine
cap_win <- cap_win %>% 
  bind_rows(cap_win_2030) %>%
  bind_rows(cap_win_2040)
# Note: the existing capacity will be linearly interpolated from 2020 to 2040
# cap = 0 after 2040

# load repository with wind technologies
EWIN <- ideea_modules$techs$EWIN
class(EWIN) # repository with wind techs
names(EWIN@data)

# update base-year technology with preexisting capacity
EWIN@data$EWIN_2020 <- EWIN@data$EWIN_2020 %>%
  update(stock = data.frame(
    region = cap_win$region,
    year = cap_win$year,
    stock = cap_win$capacity_mw / 1e3 # in GW
  ))

# add weather factor name and parameter for each technology
EWIN@data <- lapply(EWIN@data, function(tech) update(
  tech,
  weather = data.frame(weather = "WWIN", waf.fx = 1),
  input = list(comm = "REN", combustion = 0)
))
names(EWIN@data)
EWIN@data$EWIN_2020@weather

EWIN@data$EWIN_2020 %>% draw()
```

### Hydro
existing - from WRI dataset, no new?
```{r}
cap_hyd <- cap %>%
  filter(grepl("Hydro", primary_fuel)) %>%
  filter(!offshore) %>%
  mutate(year = 2020, .before = 1)

# assume retirement of 2020 capacity by 2030
cap_hyd_2030 <- cap_hyd %>%
  mutate(year = 2030, capacity_mw = 1 * capacity_mw)

# assume retirement of 2020 capacity by 2030
cap_hyd_2040 <- cap_hyd %>%
  mutate(year = 2060, capacity_mw = 1 * capacity_mw)

# combine
cap_hyd <- cap_hyd %>% 
  bind_rows(cap_hyd_2030) %>%
  bind_rows(cap_hyd_2040)
# Note: the existing capacity will be linearly interpolated from 2020 to 2040
# cap = 0 after 2040

# load base-year technology 
EHYD_2020 <- ideea_modules$techs$EHYD@data$EHYD_2020
class(EHYD_2020) 

# update base-year technology with preexisting capacity
EHYD_2020 <- EHYD_2020 %>%
  update(stock = data.frame(
    region = cap_hyd$region,
    year = cap_hyd$year,
    stock = cap_hyd$capacity_mw / 1e3 # in GW
  ))

EHYD <- ideea_modules$techs$EHYD %>% #  
  add(EHYD_2020, overwrite = T) # add tech with existing capacity
names(EHYD@data)

draw(EHYD@data$EHYD_2020)

```

### Nuclear
```{r}
cap_nuc <- cap %>%
  filter(grepl("Nuclear", primary_fuel)) %>%
  filter(!offshore) %>%
  mutate(year = 2020, .before = 1)

# assume retirement of 2020 capacity by 2060
cap_nuc_2030 <- cap_nuc %>%
  mutate(year = 2030, capacity_mw = 1 * capacity_mw)

# assume retirement of 2020 capacity by 2030
cap_nuc_LAST <- cap_nuc %>%
  mutate(year = 2060, capacity_mw = 1 * capacity_mw)

# combine
cap_nuc <- cap_nuc %>% 
  bind_rows(cap_nuc_2030) %>%
  bind_rows(cap_nuc_LAST)
# Note: the existing capacity will be linearly interpolated from 2020 to 2040
# cap = 0 after 2040

# load base-year technology 
ENUC_2020 <- ideea_modules$techs$ENUC@data$ENUC_2020
class(ENUC_2020) 

# update base-year technology with preexisting capacity
ENUC_2020 <- ENUC_2020 %>%
  update(stock = data.frame(
    region = cap_nuc$region,
    year = cap_nuc$year,
    stock = cap_nuc$capacity_mw / 1e3 # in GW
  ))

ENUC <- ideea_modules$techs$ENUC %>% #  
  add(ENUC_2020, overwrite = T) # add tech with existing capacity
names(ENUC@data)

draw(ENUC@data$ENUC_2020)

```

### CCS
tbc
```{r}

```

### Nuclear
```{r}
cap_nuc <- cap %>%
  filter(grepl("Nuclear", primary_fuel)) %>%
  filter(!offshore) %>%
  mutate(year = 2020, .before = 1)

# assume retirement of 2020 capacity by 2030
cap_nuc_2030 <- cap_nuc %>%
  mutate(year = 2030, capacity_mw = 1 * capacity_mw)

# assume retirement of 2020 capacity by 2030
cap_nuc_2040 <- cap_nuc %>%
  mutate(year = 2060, capacity_mw = 1 * capacity_mw)

# combine
cap_nuc <- cap_nuc %>% 
  bind_rows(cap_nuc_2030) %>%
  bind_rows(cap_nuc_2040)
# Note: the existing capacity will be linearly interpolated from 2020 to 2040
# cap = 0 after 2040

# load coal technology (assume Super-critical for all existing capacity)
ENUC_2020 <- ideea_modules$techs$ENUC@data$ENUC_2020
class(ENUC_2020) 

# update base-year technology with preexisting capacity
ENUC_2020 <- ENUC_2020 %>%
  update(stock = data.frame(
    region = cap_nuc$region,
    year = cap_nuc$year,
    stock = cap_nuc$capacity_mw / 1e3 # in GW
  ))

# load most advanced coal techs for new investment
ENUC <- ideea_modules$techs$ENUC %>% # ultra-super-critical 
  add(ENUC_2020, overwrite = T) # add tech with existing capacity
names(ENUC@data)

draw(ENUC@data$ENUC_2020)
```

### Bio energy
bio-reactors, agricultural waste/residue 
```{r}
cap_bio <- cap %>%
  filter(grepl("Biomass", primary_fuel)) %>%
  filter(!offshore) %>%
  mutate(year = 2020, .before = 1)

# assume retirement of 2020 capacity by 2030
cap_bio_2030 <- cap_bio %>%
  mutate(year = 2030, capacity_mw = 1 * capacity_mw)

# assume retirement of 2020 capacity by 2030
cap_bio_2040 <- cap_bio %>%
  mutate(year = 2060, capacity_mw = 1 * capacity_mw)

# combine
cap_bio <- cap_bio %>% 
  bind_rows(cap_bio_2030) %>%
  bind_rows(cap_bio_2040)
# Note: the existing capacity will be linearly interpolated from 2020 to 2040
# cap = 0 after 2040

# load base-year technology 
EBIO_2020 <- ideea_modules$techs$EBIO@data$EBIO_2020
class(EBIO_2020) 

# update base-year technology with preexisting capacity
EBIO_2020 <- EBIO_2020 %>%
  update(stock = data.frame(
    region = cap_bio$region,
    year = cap_bio$year,
    stock = cap_bio$capacity_mw / 1e3 # in GW
  ))

EBIO <- ideea_modules$techs$EBIO %>% #  
  add(EBIO_2020, overwrite = T) # add tech with existing capacity
names(EBIO@data)

draw(EBIO@data$EBIO_2020)

```

## Energy storage
```{r}
STG_BTR <- ideea_modules$techs$STG_BTR

```


## Transmission
```{r}
transmission_existing <- ideea_data$transmission_5x5 %>%
  filter(scenario == "current", !is.na(MW)) %>%
  rename(dst = destination) %>%
  filter(!duplicated(MW)) # temporary solution

gis_mainland_sf <- filter(gis_sf, offshore == F, mainland == T)
points_coord <- st_centroid(gis_mainland_sf) %>%
  st_coordinates() %>%
  as.data.frame() %>%
  cbind(data.frame(region = gis_mainland_sf$region)) %>%
  rename(lon = X, lat = Y)

a <- transmission_existing %>%
  # filter(scenario == "current") %>%
  left_join(points_coord, by = c("region" = "region")) %>%
  left_join(points_coord, by = c("dst" = "region")) %>%
  mutate(GW = MW/1000)

ggplot() +
  geom_sf(data = gis_sf, fill = "wheat") +
  geom_segment(aes(x = lon.x, y = lat.x, xend = lon.y, yend = lat.y, 
                   linewidth = GW),
               color = "dodgerblue", lineend = "round",
               data = filter(a, scenario == "current")) +
  geom_segment(aes(x = lon.x, y = lat.x, xend = lon.y, yend = lat.y),
               color = alpha("white", .5), lineend = "round",
               data = filter(a, scenario == "new")) +
  geom_point(aes(lon, lat), data = points_coord, color = "red") +
  labs(x = "", y = "") +
  theme_bw()
ggsave("tmp/network_5x5.png", width = 5, height = 5)
  
network <- transmission_existing %>%
  filter(scenario == "current") %>%
  mutate(
    trd_name = paste("TRD", region, dst, sep = "_")
  ) %>%
  rename(
    region.x = region,
    region.y = dst
  )

repo_transmission_existing <- newRepository("transmission_existing")
if (nrow(network) > 0) {
  for (i in 1:nrow(network)) {
    trd <- newTrade(
      name = network$trd_name[i],
      # description = network$lines_type[i],
      commodity = "ELC",
      routes = data.frame(
        src = c(network$region.x[i], network$region.y[i]),
        dst = c(network$region.y[i], network$region.x[i])
      ),
      trade = data.frame(
        src = c(network$region.x[i], network$region.y[i]),
        dst = c(network$region.y[i], network$region.x[i]),
        teff = rep(1 - .1, 2)
      ),
      capacityVariable = T,
      # invcost = data.frame(
      #   # costs can be assigned to one of the connected region or both
      #   # here we split the costs, 50% for each region
      #   region = c(network$region.x[i], network$region.y[i]),
      #   invcost = rep(1000, 2) * 2 # olife == 2
      # ),
      # olife = 50, # doubled annualized invcost for consistency 
      end = base_year - 100,
      stock = data.frame(
        year = c(2020, 2070),
        stock = c(network$MW[i], network$MW[i])
      ),
      cap2act = 24*365
    )
    repo_transmission_existing <- add(repo_transmission_existing, trd)
    rm(trd)
  }
}
names(repo_transmission_existing)

transmission_new <- ideea_data$transmission_5x5 %>%
  filter(scenario == "new", !is.na(MW)) %>%
  rename(dst = destination) %>%
  filter(!duplicated(MW)) # temporary solution

repo_transmission_new <- newRepository("transmission_new")
if (nrow(network) > 0) {
  for (i in 1:nrow(network)) {
    trd <- newTrade(
      name = str_replace(network$trd_name[i], "TRD_", "TRN_"),
      # description = network$lines_type[i],
      commodity = "ELC",
      routes = data.frame(
        src = c(network$region.x[i], network$region.y[i]),
        dst = c(network$region.y[i], network$region.x[i])
      ),
      trade = data.frame(
        src = c(network$region.x[i], network$region.y[i]),
        dst = c(network$region.y[i], network$region.x[i]),
        teff = rep(1 - .1, 2)
      ),
      capacityVariable = T,
      invcost = data.frame(
        # costs can be assigned to one of the connected region or both
        # here we split the costs, 50% for each region
        region = c(network$region.x[i], network$region.y[i]),
        invcost = rep(1000, 2) / 2 # olife == 2
      ),
      olife = 50, # doubled annualized invcost for consistency 
      start = base_year - 10,
      # stock = data.frame(
      #   # year = 
      #   stock = network$MW[i]
      # ),
      cap2act = 24*365
    )
    repo_transmission_new <- add(repo_transmission_new, trd)
    rm(trd)
  }
}
names(repo_transmission_new)

```

## Unserved load
```{r}
UNSERVED <- newImport(
  name = "UNSERVED",
  description = "Unserved load, ~ 10 cr.INR/GWh, equal 100 INR/kWh",
  commodity = "ELC",
  unit = "GWh",
  imp = list(price = 10)
)

```



## Policies
  
```{r}
CO2_CAP <- newConstraint("CO2_CAP",
    eq = "<=", 
    rhs = list(year = c(2020, 2055), 
               rhs = c(300000, 0)),
    for.each = list(year = NULL), # Cap total emission
    variable = list(variable = "vBalance", 
                    for.sum = list(comm = "CO2",
                                   slice = NULL,
                                   region = NULL))
  )
```

```{r}

NO_NEW_HYDRO <- newConstraint(
  name = "NO_NEW_HYDRO",
  # description = "Constraint on new Hydro plants constraction",
  eq = "<=",
  for.each = list(year = c(2020, 2055)),
  variable = list(
    variable = "vTechNewCap",
    for.sum = list(
      tech = c("EHYD_2030", "EHYD_2040", "EHYD_2050", "EHYD_2020"),
      region = NULL
    )
  ),
  rhs = list(
    year = c(2020, 2055),
    rhs = c(0, 0)
  )
)

NO_NEW_NUCLEAR <- newConstraint(
  name = "NO_NEW_NUCLEAR",
  # description = "Constraint on new Hydro plants constraction",
  eq = "<=",
  for.each = list(year = c(2020, 2055)),
  variable = list(
    variable = "vTechNewCap",
    for.sum = list(
      tech = c("ENUC_2030", "ENUC_2040", "ENUC_2050", "ENUC_2020"),
      region = NULL
    )
  ),
  rhs = list(
    year = c(2020, 2055),
    rhs = c(0, 0)
  )
)

dput(names(EGAS))
CT_EGAS <- newConstraint(
  name = "CT_EGAS",
  # description = "Constraint on new Hydro plants constraction",
  eq = "<=",
  for.each = list(year = c(2020, 2055)),
  variable = list(
    variable = "vTechNewCap",
    for.sum = list(
      tech = dput(names(EGAS)),
      region = NULL
    )
  ),
  rhs = list(
    year = c(2020, 2055),
    rhs = c(5, 5)
  )
)
# limit on hydro construction

```


## Model
```{r}
repo <- newRepository("repo") %>%
  add(
    # commodities
    # ELC, COA, CO2, NOX, SOX,
    repo_comm,
    # supply & import
    repo_supply,
    # Generating technologies
    ECOA,
    # EGAS,
    # CT_EGAS,
    # ENUC,
    ENUC,
    # NO_NEW_NUCLEAR,
    # EHYD,
    EHYD,
    # NO_NEW_HYDRO, # limit on hydro construction
    ESOL, WSOL,
    EWIN, WWIN,
    # STG_BTR_2020,
    # STG_BTR_2030,
    STG_BTR,
    # transmission
    # repo_transmission_existing,
    repo_transmission_new,
    # unserved load
    UNSERVED,
    # demand
    DEMELC_BY,
    DEMELC_2X
  )
  # fract_year_adj_repo(subset_hours = SUBSET_HOURS)

length(repo)
print(repo)
names(repo)
summary.repository(repo)

# model-class object 
mod <- newModel(
  name = 'IDEEA_r7_base', 
  description = "IDEEA example model",
  ## in case of infeasibility, `dummy` variables can be added
  # debug = data.frame(#comm = "ELC",
  #                    dummyImport = 1e6,
  #                    dummyExport = 1e6),
  region = unique(gis_sf$region),
  discount = 0.05,
  slice = tsl_levels,
  yearFraction = list(fraction = FRACT_YEAR),
  # slice = tsl_sets,
  repository = repo)

# Check the model time-slices
mod@sysInfo@slice@levels %>% as.data.table()

# Set milestone-years
# mdl <- setMilestoneYears(mdl, start = 2015, interval = c(1, 2, 5, 6, 7, rep(10, 3)))
# mdl <- setMilestoneYears(mdl, start = baseYear, interval = c(1, diff(modYears)))
mod <- setMilestoneYears(mod, 
                         start = 2030,
                         # start = base_year,
                         interval = c(1))
                         # interval = c(1, 5, 10))
                         # interval = c(1, 3, 4, 5, 5, 5, 5, 5))
                         # interval = c(1, 5, 10, 10, 10))
mod@sysInfo@milestone # check
```


## Scenario
```{r scen_gams, eval=FALSE, echo=T}
# solver_gams$solver <- "Cbc"
# mod <- add(mod, CO2_CAP)

scen_gms <- solve(mod, 
                  name = "IDEEA_test", 
                  solver = solver_gams,
                  # solver = GAMS_csv_cbc,
                  # solver = GAMS_csv_cplex,
                  tmp.dir = file.path("tmp", "GAMS"))

if (F) {
  scen <- interpolate(mod, name = "IDEEA_toy", trim = F)
  
  scen <- write(scen, solver = GAMS_custom, 
                tmp.dir = file.path("tmp", "GAMS2"))
  
  scen@misc$tmp.dir 
  solve(scen, wait = F)
  scen_gms <- read(scen)
}
summary(scen_gms)

```

```{r scen_julia, eval=FALSE,echo=T}
scen_jl <- solve(mod, 
                 name = "IDEEA_test", solver = solver_jl, 
                 tmp.dir = file.path("tmp", "JuMP"))
summary(scen_jl)

if (F) {
  # alternative step-by-step, manual (for lare models)
  scen <- interpolate(mod, name = "IDEEA_toy", trim = F)
  scen_jl <- write(scen, 
   solver = solver_jl,
   tmp.dir = file.path("tmp", "JuMP")
  )
  scen_jl@misc$tmp.dir 
  scen_jl <- solve(scen_jl, wait = T)
  scen_jl <- read(scen_jl)
  summary(scen_jl)
}
summary(scen_jl)

```

```{r scen_python, eval=FALSE, echo=T}

scen <- interpolate(mod, 
                    name = "IDEEA_toy", trim = F)
scen_py <- write(scen, 
 solver = solver_py,
 tmp.dir = file.path("tmp", "Pyomo")
)
scen_py@misc$tmp.dir
scen_py <- solve(scen_py, wait = T)
scen_py <- read(scen_py)
summary(scen_py)
```

```{r, eval=FALSE}
scen_glpk <- solve(mod, name = "IDEEA_r7_glpk", 
                   solver = "GLPK", 
                  tmp.dir = file.path("tmp", "GLPK"))
summary(scen_glpk)

```

## Results
```{r, eval=T}
drop_vintage <- function(x) {
  sub("_[0-9]+$", "", x)
}

drop_process_vintage <- function(x) {
  x %>% mutate(process = drop_vintage(process))
}
```

class scenario
```{r}
# rename solved scenario (one of) 
scen <- scen_gms
# scen <- scen_jl
# scen <- scen_py
class(scen)
summary(scen)
scen@name
slotNames(scen)
scen
```


Use `getData` function to get a variable or parameter of interest.
Interpolated parameters 
The full list of all the model variables can be found here:  
https://energyrt.github.io/book/variables.html 
list of all parameters: 
https://energyrt.github.io/book/parameters.html 
The interpolated parameters are stored in scenario class, `scen@modInp` slot, variables (solution) in `scen@modOut`.  
```{r, eval=FALSE}
# get total capacity of technologies.
vTechCap <- getData(scen, 
                    "vTechCap", merge = T, drop.zeros = T,
                    process = T) %>%
  drop_process_vintage() %>%
  group_by(scenario, process, region, year) %>%
  summarise(value = sum(value), .groups = "drop")
vTechCap

# get all new installations (technologies, storage, trade)
vNewCap <-
  getData(scen, name_ = "New", merge = T, 
          process = T, drop.zeros = T) %>%
  drop_process_vintage() %>%
  group_by(scenario, process, region, year) %>%
  summarise(value = sum(value), .groups = "drop")
vNewCap

# the same - all new capacity without merging, by variable
  getData(scen, name_ = "NewCap", merge = F, drop.zeros = T)
# everything related to trade
  getData(scen, name_ = "Trade", merge = F, drop.zeros = T)
# new trade routes (transmission)
  getData(scen, name_ = "TradeNew", merge = F, drop.zeros = T)
# all related to storage
  getData(scen, name_ = "Storage", merge = F, drop.zeros = T, 
          parameters = F, digits = 3)
# Import from the RoW
  getData(scen, name = "vImportRow", parameters = F,  merge = T, drop.zeros = T) %>% as.data.table()

# plot capacity
vTechCap_sf <- gis_sf %>% 
  filter(!offshore) %>%
  left_join(vTechCap, by = c(reg_off = "region"))
  
ggplot(vTechCap_sf) +
  geom_sf(aes(fill = value), na.rm = F) +
  scale_fill_viridis_c(option = "H", name = "GW") +
  facet_grid(year~process) +
  theme_bw()

# plot new capacity
vNewCap_sf <- gis_sf %>% 
  filter(!offshore) %>%
  left_join(select(vNewCap, -any_of("name")), 
            by = c(reg_off = "region")) %>%
  filter(!is.na(value))
st_drop_geometry(vNewCap_sf) %>% as.data.table()

ggplot(vNewCap_sf) +
  geom_sf(data = gis_sf, na.rm = F) +
  geom_sf(aes(fill = value), na.rm = F) +
  scale_fill_viridis_c(option = "H", name = "GW") +
  facet_grid(process~year) +
  theme_bw()

```

Emissions
```{r}
getData(scen,"vBalance", comm = "CO2", merge = T) %>%
  group_by(scenario, year) %>% summarise(CO2 = sum(value))

getData(scen, parameters = T, variables = F, comm = "CO2", 
        merge = F)
scen@modInp@parameters$pCnsRhsCO2_CAP

getData(scen, name_ = "CO2_CAP")

```

Generation by technology type
```{r}
vTechOut <- getData(scen, 
                    "vTechOut", comm = "ELC", 
                    merge = T, drop.zeros = T,
                    process = T) %>%
  drop_process_vintage %>%
  group_by(scenario, name, process, year) %>%
  summarize(
    GWh = sum(value, na.rm = T),
    .groups = "drop"
  )

ggplot(vTechOut) +
  geom_bar(aes(year, GWh, fill = process), stat = "identity") 
  # facet_wrap(~region)
  
  
getData(scen, "pDemand", merge = T) %>%
  group_by(scenario, dem, year) %>%
  summarise(value = sum(value))

```

All new capacity incl. trade
```{r}
vCap <- getData(scen, name_ = "New",
                    # "vTechCap", 
                    merge = T, drop.zeros = T,
                    process = T) %>%
  drop_process_vintage %>%
  group_by(scenario, name, process, year) %>%
  summarize(
    GW = sum(value, na.rm = T),
    .groups = "drop"
  )

ggplot(vCap) +
  geom_bar(aes(year, GW, fill = process), stat = "identity") 
  # facet_wrap(~region)
  
# demand(s)  
getData(scen, "pDemand", merge = T) %>%
  group_by(scenario, dem, year) %>%
  summarise(value = sum(value))

```

compare gms, jl, py, glpl
```{r, eval=FALSE}
sns <- list(
  GAMS = scen_gms, 
  Julia = scen_jl,
  Pyomo = scen_py
  ,GLPK = scen_glpk
  )

getData(sns, "vObjective", merge = T) %>% as.data.table()
# save(sns, file = "tmp/four_identical_solutions.RData")

getData(sns, name_ = "NewCap", merge = T, process = T,
        digits = 2, drop.zeros = T) %>% 
  pivot_wider(names_from = "scenario") %>%
  select(-name) %>%
  as.data.table()
```

tbc.
