---
title: "Setup"
author: '@ideea'
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation check-list

### 1. **R** and **RStudio** 

- <https://cran.rstudio.com/> 
- <https://posit.co/download/rstudio-desktop/> 

### 2. Optimization software

At least one of the following optimization, algebraic modeling software is required to solve the model:   
- **GAMS** (proprietary, license is required) 
- **Python** with *Pyomo* and solver libraries (*free*, can be connected to proprietary solvers such as CLPEX, GUROBI)  
- **Julia** with JuMP and solver libraries (*free*, can be connected to proprietary solvers such as CLPEX, GUROBI)  
- **GLPK**, standalone solver and modeling language (free, *for small models only*) 

The software is listed in descending order of time-performance to get the model/scenario solved. Large models with full sets of time-slices (like hourly) require powerful solvers like CPLEX or GUROBI. Both are available for free for academia.

See details: ...[link]

### 3. R libraries. 
```{r}
remotes::install_github("energyRt/energyRt", ref = "dev")
remotes::install_github("energyRt/timeslices")
remotes::install_github("ideea-model/IDEEA")

```


### PATH to optimization software

The optimization software should be "visible" from R environment. The straightforward way to achieve this is to add the software directory to the system PATH.

```{r, eval=FALSE}
# check path to your software
Sys.getenv("PATH") 
```

If the path to your optimization software is not in the system PATH variables or different version/environment is planned to use in this project, it is recommended specify the path in `cmdline` of solver parameters.\
An example with alternative choices of LP solvers (change paths according to the libraries location on your system):

```{r, eval=FALSE}
solver_gams <- list(
  cmdline = "/Applications/GAMS25.1/sysdir/gams energyRt.gms", # MacOS/Linux
  # cmdline = "C:/GAMS/win64/32.2/gams.exe energyRt.gms", # Windows
  lang = "GAMS",
  solver = "cplex"
  # solver = "cbc"
)

solver_jl <- list(
  cmdline = "/Applications/Julia-1.6.app/Contents/Resources/julia/bin/julia energyRt.jl", # MacOS/Linux
  lang = "JuMP",  
  # solver = "Cbc"
  solver = "HiGHS"
  # solver = "cplex"
  # solver = "gurobi"
)

# to list all available conda environments: 
reticulate::conda_list()

solver_py <- list(
  cmdline = "/Users/olmba/Library/r-miniconda-arm64/envs/switch/bin/python energyRt.py", # MacOS/Linux
  lang = "PYOMO",
  export_format = "SQLite",
  solver = "cbc"
  # solver = "cplex"
  # solver = "glpk"
  # solver = "highs"
)

solver_glpk <- "GLPK" # additional specification for this optimization software is not available, the path to the library must be specified in the global `PATH` variables. 

```

The paths are specific for every system, here are couple examples for Windows and Mac OS systems.
